# (PART) Linear Algebra in R {-}

# Class Examples

## Day 3: Wed Sep 8

Remember that to use RREF we must include the `pracma` package.
```{r}
require('pracma')
```

Our goal is to solve the following system of equations:
$$ 
\left\{\begin{array}{rrrrrrrrrr}
x_1 &+& 2 x_2 &+&x_3 &+&  x_4 & = & 4 \\
x_1 &+& 2x_2 &+& -x_3 &+& -3x_4& =& 6 \\
 && x_2 &+&  x_3 &+&  x_4& =& 0 \\
-x_1&+&  x_2 &+& -x_3&+&  -4x_4& =  &-1\\
\end{array}
\right\}
$$

We enter the augmented matrix (and echo it back):
```{r,echo=TRUE}
A = cbind(c(1,1,0,-1),c(2,2,1,1),c(1,-1,1,-1),c(1,-3,1,-4),c(4,6,0,-1))
A
```

And then we row reduce it using `rref`:
```{r,echo=TRUE}
rref(A)
```

The corresponding reduced set of equations is
$$ 
\left\{\begin{array}{rrrrrrrrrr}
x_1 &&  && &+&  x_4 & = & 3 \\
 && x_2 &&  &-& x_4& =& 1 \\
 &&  &+&  x_3 &+&  2x_4& =& -1 \\
\end{array}
\right\}
$$

The **general** solution to this system of equations is
$$
\begin{align}
x_1 &= 3 - x_4 \\
x_2 &= 1 - x_5 \\
x_3 &= -1 - 2 x_4 \\
x_4 &= free
\end{align}
$$

## Day 4: Fri Sep 10

**Question** Solve the following matrix equation using R:
$$
\begin{bmatrix}
 1 & -2 & 1 \\
 1 & 0 & 1 \\
 -1 & 1 & ~0~ \\
 2 & 1 & 1  \\
\end{bmatrix}
\begin{bmatrix} x_1 \\ x_2 \\ x_3 \end{bmatrix}
=
\begin{bmatrix} 13 \\ 7 \\ -5 \\ 6  \end{bmatrix}
$$

Here we enter A and b separately to illustrate (below) how to augment a matrix with a vector.
```{r,echo=TRUE}
A = cbind(c(1,1,-1,2),c(-2,0,1,1),c(1,1,0,1))
b = c(13,7,-5,6) # notice that b is a vector and not a matrix.
```

Let's echo them back to see what they look like
```{r}
A
b
```
Now I will augment A with b and call it Ab. The syntax is nice:

```{r,echo=TRUE}
Ab = cbind(A,b)
Ab
```

And then row reduce.

```{r,echo=TRUE}
rref(Ab)
```
This tells me that there is a **unique** solution to Ax = b, and it is 
$$
\begin{bmatrix} x_1 \\ x_2 \\ x_3  \end{bmatrix} = 
\begin{bmatrix} 2 \\ -3 \\ 5  \end{bmatrix}.
$$
**Question 2** Can we find a vector d for which A x = d has no solution?

```{r,echo=TRUE}
A = cbind(c(1,1,-1,2),c(-2,0,1,1),c(1,1,0,1))
d = c(1,1,-1,2) # notice that b is a vector and not a matrix.
Ad = cbind(A,d)
rref(Ad)
```

## Day 5: Mon Sep 13

<!-- Give the parametric solution to $A x = b$ if  -->
<!-- $$ -->
<!-- A=\left[ -->
<!-- \begin{array}{ccccc} -->
<!--  1 & 1 & 1 & -1  \\ -->
<!--  1 & 0 & 3 & -1  \\ -->
<!--  -1 & 1 & -5 & 1  \\ -->
<!--  1 & 2 & -1 & 1  \\ -->
<!-- \end{array} -->
<!-- \right] \qquad \hbox{and}\qquad  b = \begin{bmatrix} 2 \\ 4 \\ -6 \\ 4 \end{bmatrix}. \hskip.5in -->
<!-- $$ -->
<!-- ```{r,echo=TRUE} -->
<!-- A = cbind(c(1,1,-1,1),c(1,0,1,2),c(1,3,-5,-1),c(-1,-1,1,1)) -->
<!-- b = c(2,4,-6,4) -->
<!-- Ab = cbind(A,b) -->
<!-- A -->
<!-- Ab -->
<!-- rref(Ab) -->
<!-- ``` -->
<!-- a. Give the solution this $Ax=b$ problem in *parametric* form: -->

<!-- $$ -->
<!-- \begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \end{bmatrix} -->
<!-- =  -->
<!-- \begin{bmatrix}  6 \\ -2 \\ 0 \\ 2 \end{bmatrix} -->
<!-- + s \begin{bmatrix}  -3 \\ 2 \\ 1 \\ 0 \end{bmatrix} -->
<!-- $$ -->

<!-- b. Let's make two different solutions. One with s = 0 and one with s = 1. They are -->

<!-- ```{r,echo=TRUE} -->
<!-- x1 = c(6,-2,0,2) -->
<!-- x2 = c(3,0,1,2) -->
<!-- ``` -->

<!-- And we check to see that they are solutions. Observe that we use the unfortunate notation `%*%` to multply a matrix times a vector, but also that we get b in both cases: -->
<!-- ```{r,echo=TRUE} -->
<!-- A %*% x1 -->
<!-- A %*% x2 -->
<!-- ``` -->
<!-- Notice that we get b for both. -->

<!-- c. Now we look at the difference. It is a solution to the *homogeneous* equations A x = 0 as we see here: -->
<!-- ```{r,echo=TRUE} -->
<!-- xh = x1 - x2 -->
<!-- A %*% xh -->
<!-- ``` -->

<!-- d. Are any of these vectors solutions to Ax=b? -->

<!-- ```{r,echo=TRUE} -->
<!-- A %*% (x1 + x2) -->
<!-- ``` -->
<!-- No! (this is 2b) -->

<!-- ```{r,echo=TRUE} -->
<!-- A %*% (1/2*x1 + 1/2* x2) -->
<!-- ``` -->

<!-- Yes! This is b -->

<!-- ```{r,echo=TRUE} -->
<!-- A %*% (x1 + 2021 * xh) -->
<!-- ``` -->
<!-- Yes! This is b. -->