
#  Network Centralities


[Download this Rmd file](https://github.com/Tom-Halverson/math236_mod5/blob/main/NetworkCentrality.Rmd). You can cut and paste it into an Rmd file.


In this example, we will use a package called `igraph`. To install it, you need to go to the packages window (bottom right), choose `install`, and search for and install `igraph` from the packages window.

```{r,message=FALSE, echo=TRUE}
library(igraph)
```

## Graphs and Networks

Graphs consists of **vertices** and the **edges** between them. These edges are used to model connections in a wide array of applications, including but not limited to, physical, biological, social, and information networks. 

### Adjacency Matrices

Matrices are used to represent graphs and networks in a very direct way: we place a 1 in position $(i,j)$ of the **adjacency matrix** $A$ of the graph $G$, if there is an edge from vertex $i$ to vertex $j$ in $G$. Here is the adjacency matrix we will use today. 
```{r, echo=TRUE}
A = rbind( c(0,1,0,1,0,0,0,0,1,0,0,0),c(1,0,1,1,1,0,1,0,0,0,0,0),c(0,1,0,0,1,0,0,0,0,0,0,0), c(1,1,0,0,0,1,0,1,0,0,0,0), c(0,1,1,0,0,0,1,1,0,0,0,1), c(0,0,0,1,0,0,1,0,0,0,0,0), c(0,1,0,0,1,1,0,1,0,0,0,0), c(0,0,0,1,1,0,1,0,0,1,1,0), c(1,0,0,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,1,0,0,0,0), c(0,0,0,0,0,0,0,1,0,0,0,0), c(0,0,0,0,1,0,0,0,0,0,0,0))
A
```


We make a graph from the adjacency matrix using the code below. Observe that there is an edge from vertex $i$ to vertex $j$ if and only if there is a 1 in position $(i,j)$ in the matrix.
```{r, network1, fig.height=6, fig.width=6, echo=TRUE}
g=graph_from_adjacency_matrix(A,mode='undirected')
plot(g, vertex.color='tan1', vertex.frame.color="dodgerblue")
```


This network is the route map of a small airline. Here we will add the locations of the vertices, add vertex labels, and change the vertex size:

```{r, network2, fig.height=6, fig.width=6, echo=FALSE}
locations = rbind(c(20,0),c(-10,0),c(11,7),c(10,15),c(3,12),c(25,10),c(-10,10),c(-12,15),c(20,6),c(-15,12),c(12,4),c(25,13))
airports = c("ATL","LAX","ORD","MSP","DEN","JFK","SFO","SEA","PHL","PDX","MDW","LGA")
V(g)$label = airports
plot(g,vertex.size=20, layout=locations, vertex.label.cex=0.85, vertex.color='tan1', vertex.frame.color="dodgerblue")
```


## Degree Centrality

If we are considering placing an office in one of our airport locations, we may want to chose the most **central** hub for that office. It turns out that there are many interesting **centrality measures** for networks. We will talk about two of them today.

The simplest measure centrality is the **degree** of the vertex, or the number of edges connected to that vertex. We calculate the degree centralities from the adjacency matrix as follows. 

1. Make a vector $\mathsf{v}$ of all 1's and multiply $\mathsf{d} = A\mathsf{v}$ to get the vector of degrees.
1. Divide the vector $\mathsf{p}$ by the sum of its entries to get the **degree** proportions. The result is a **normalized vector** $\mathsf{p}$ whose entries sum to 1. Each entry of vector $\mathsf{p}$ represents to *proportion* of edges incident with the corresponding vertex.

```{r, echo=TRUE}
v=rep(1,nrow(A)) # all 1s vector
d = A %*% v  # degrees
p=d/sum(d)   # proportion of degrees
Z = cbind(d,p) # show d and p together side-by-side in a matrix
rownames(Z) = airports
colnames(Z) = c("deg","proportion")
ii=order(d,decreasing=TRUE) # sort from largest to smallest
Z = Z[ii,]
Z
```


Now we create a *data visualization* that plots the  size of each vertex according to the vector $p$. The larger vertices have more edges connected to them. 

```{r, network7, fig.height=6, fig.width=6, echo=TRUE}
plot(g, layout=locations, vertex.size=250*p,vertex.label.cex=0.65, vertex.color='tan1', vertex.frame.color="dodgerblue")
```


## Gould's Index

Gould's Index is a measure of centrality that uses the dominant eigenvector of a matrix. It was introduced by geographer P. R. Gould in 1967 to analyze the geographical features on maps. We will build up Gould's Index step-by-step so that we can understand what it measures. 

### Step 1: Add Layovers

The first step is typically to add the identity matrix to the adjancency matrix $A$ to get a new matrix 
$$
B = A + I.
$$ 
The $n \times n$ identity matrix in `R` is obtained by using `diag(n)`. Adding the identity gives a connection from a vertex to itself. This **loop edge** corresponds to staying at the current city during a layover. 

```{r, echo=TRUE}
(B = A + diag(nrow(A)))
```


```{r, networkB, fig.height=6, fig.width=6, echo=TRUE}
g2=graph_from_adjacency_matrix(B,mode='undirected')
V(g2)$label = airports
plot(g2, layout=locations, vertex.color='tan1', vertex.frame.color="dodgerblue")
```


### Step 2: Dynamical System

Starting with the all 1's vector $\mathsf{v}_0$ create the dynamical system
$$
\mathsf{v}_0, \quad  
\mathsf{v}_1 = B \mathsf{v}_0,  \quad 
\mathsf{v}_2 = B \mathsf{v}_1,  \quad
\mathsf{v}_3 = B \mathsf{v}_2,  \quad \ldots , \quad
\mathsf{v}_n  = B \mathsf{v}_{n-1}.
$$
Here we calculate $\mathsf{v}_1, \ldots, \mathsf{v}_{10}$ using a loop:
```{r, echo=TRUE}
N = 10 
X = matrix(0,nrow=nrow(B),ncol=N+1) # make a a table of 0s
X[,1] = rep(1,nrow(B))              # put v0 in first column
for (i in 1:N) {                  # loop N times
   X[,i+1] = B %*%  X[,i]   # apply B to the ith column and make it the (i+1)st column      
}
rownames(X) = airports
colnames(X) = 0:10
X
```

**Discuss with your group:** Each of the entries of the vector $\mathsf{v}_{t}$ in the columns of the table above corresponds to "a trip of length $t$." What kinds of trips do the entries of $\mathsf{v}_{t}$ count? To figure this out, compare the table of vectors with the picture of the network with layovers. Figure out the maning of the $t=1$ column, then the $t = 2$ column, and so on. Why does the rule $\mathsf{v}_t = B \mathsf{v}_{t-1}$ leads to this result?


**Normalize**: These numbers get big fast! Let's normalize by dividing by the sum each time. By doing this, the vectors will always be proportions which sum to 1. See the table below. What do the entries in this table tell us?

```{r, echo=TRUE}
N = 10 
X = matrix(0,nrow=nrow(B),ncol=N+1)
X[,1] = rep(1,nrow(B))
for (i in 2:(N+1)) {  
   X[,i] = B %*%  X[,i-1]
   X[,i] = X[,i]/sum(X[,i])
}
rownames(X) = airports
colnames(X) = 0:10
X
```

### Step 4: Eigen-analysis 

We see that the vectors are converging to a common direction, and 
we know that dynamical systems converge to the **dominant eigenvector** (if there is one). We can see below that there is a dominant eigenvector in this case. 
```{r, echo=TRUE}
eigen(B)
```

For an adjacency matrix $A$, the dominant eigenvector of $B + I$, scaled to sum to 1, is called **Gould's Index** of network centrality. Here we extract it, scale it to sum to 1, and we show that the dynamical system is converging to it.

```{r, echo=TRUE}
vecs = eigen(B)$vectors
gould = vecs[,1]
gould = gould/sum(gould)

# Compute the dynamical system
N = 30
X = matrix(0,nrow=nrow(B),ncol=N+1)
X[,1] = rep(1,nrow(B))/nrow(B)
for (i in 1:N) {  
   X[,i+1] = B %*%  X[,i]
   X[,i+1] = X[,i+1]/sum(X[,i+1])
}

# Display the data
Y = cbind(X[,1],X[,2],X[,3],X[,11],X[,21],gould)
rownames(Y) = airports
colnames(Y) = cbind("n=0","n=1","n=2","n=10","n=20","Gould")
Y
```


### Step 5: Visualize

Now let's plot the network with:

* the vertices sized by Gould's Index
* the labels sized by degree centrality

```{r, network6, fig.height=6, fig.width=6, echo=TRUE}
plot(g, layout=locations, vertex.size=250*gould,vertex.label.cex=8*p, 
     vertex.color='tan1',vertex.frame.color="dodgerblue" )
```

And we show the data containing Gould's Index and the Degree Centrality. We order the data using the Gould Index and then compare the two. Observe that degree centrality and Gould's Index do not always agree.
```{r, echo=TRUE}
Z = cbind(gould,p) 
rownames(Z)=airports
colnames(Z)=c('Gould', 'Degree')
ii=order(gould,decreasing=TRUE)
Z = Z[ii,]
Z
```

**Discuss with your group:** Degree centrality and Gould's Index give different rankings. Look at the table and observe that:

* LAX, DEN and SEA have the same degree centrality. However LAX and DEN have higher Gould Index than SEA.
* SFO has lower degree centrality than SEA, but higher Gould centrality! So these two centralities give different rankings.
* Why does the Gould Index value SFO more than SEA?
* Find another pair of cities where the rankings of degree centrality and Gould's Index differ. Look at the plot of the network and explain why this is the case.


### Gould Index Summary

Now that we understand what Gould's Index means, let's summarize how to find the Gould Index values for an adjacency matrix $A$.

1. Create the matrix $B = A+I$.
2. Find the dominant eigenvector $\mathbf{v}$ of $B$.
3. Normalize the values of $\mathbf{v}$ so that the entries sum to 1.




