---
title: "PS-10-2-3-template.Rmd"
author: "Your Name"
date: "11/30/2021"
output: html_document
---

```{r}
require('pracma')
```

## Problem 10.2. Projection onto subspace without orthogonal basis

**a. Fibonacci Subspace**: Insert the R-code here that (i) finds the normal equations, (ii) solves them for xhat, (iii) find the projection vhat, (iv) finds the residual, (v) shows that the residual is in $F^\perp$.

```{r}

```

Then typeset your answer here:
$$
\underbrace{\begin{bmatrix} 4\\ -1 \\ 2 \\  1 \\ -2 \end{bmatrix}}_v
= 
\underbrace{\begin{bmatrix} V \\ W \\ X \\ Y \\ Z \end{bmatrix}}_{\hat v \in F}
+
\underbrace{\begin{bmatrix} A \\ B \\ C \\ D \\ E \end{bmatrix}}_{r \in F^\perp}
$$

**b. Zero-Sum Subspace**: Insert the R-code here that (i) finds the normal equations, (ii) solves them for xhat, (iii) find the projection vhat, (iv) finds the residual, (v) shows that the residual is in $F^\perp$.

```{r}

```

Then typeset your answer here:
$$
\underbrace{\begin{bmatrix} 4\\ -1 \\ 2 \\  1 \\ -2 \end{bmatrix}}_v
= 
\underbrace{\begin{bmatrix} V \\ W \\ X \\ Y \\ Z \end{bmatrix}}_{\hat v \in Z}
+
\underbrace{\begin{bmatrix} A \\ B \\ C \\ D \\ E \end{bmatrix}}_{r \in Z^\perp}
$$

## Problem 10.3 Pseudoinverse and Projection

```{r}
(A = cbind(c(1,2,-1,-2),c(1,2,3,4),c(1,0,1,0)))
b = c(9,5,5,8)
```

a. A is not orthogonal:
```{r}

```

b. b is not in W = Col(A):
```{r}

```

c. Least-squares solution
```{r}

```

d. Residual
```{r}
```

e. Pseudoinverse
```{r}
```

f. Projection

```{r}
```


g. Projection, cont. 

```{r}
b2 = c(1,2,3,4)
```


h. Projection, cont.

```{r}
P %*% P
P
```

i. Eigenanalysis:
```{r}
eigen(P)
```
