
# Problem Set 5

* Due: Wednesday October 13 by 11:59pm CST. 
* You will get a chance to discuss some of these in class with your classmates.
* Upload your solutions to Moodle in a PDF. 
* Please feel free to **use RStudio for all row reductions.**
* You can download the [Rmd source file  for this problem set](https://github.com/Tom-Halverson/math236_f21/blob/main/PS5-problem-set-5.Rmd).



## Fibonacci Vectors

The Fibonacci vectors $F$ in $\mathbb{R}^5$ are defined below:
$$
F  = \left\{ \begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \\ x_5 \end{bmatrix} ~\Bigg\vert~ \ x_3 = x_1 + x_2, x_4 = x_2 + x_3, x_5 = x_3 + x_4 \right\} \subseteq \mathbb{R}^5.   \hskip5in
$$

a. Show that $F$ is a subspace by showing that it is closed under addition and scalar multiplication

a. Show that $F$ is a subspace by finding a matrix $A$ such that $F = Nul(A)$.

a. Show that $F$ is a subspace by writing it as the span of a set of vectors (hint: "bake" the subspace conditions into the vectors as we do in class for the zero-sum subspace).

a. Find a basis for $F$ from among your spanning vectors in the previous part. The previous part gives  you spanning. Now you need to argue that the vectors are linearly independent.

a. The dimension of a vector space is the number of vectors in a basis for the vector space. What is the dimension of $F$.


## Vectors Rescaled

If the function $T: \mathbb{R}^n \to \mathbb{R}^n$ is a linear transformation, then show that the set below is a subspace of $\mathbb{R}^n$
$$
E_{2021} = \{\ \vec{x} \in \mathbb{R}^n \mid T(\vec{x}) = 2021 \vec{x} \}.
$$
Is there anything special about 2021 in the definition? If it were replaced by another scalar, would it still be a subspace?

## Getting Into a Subspace

Let $S \subseteq \mathbb{R}^n$ be a subspace and let $\mathsf{v}, \mathsf{w} \in \mathbb{R}^n$. For  each of the following statements, either give  a specific example  or explain why it cannot happen.


a. If $\mathsf{v}$ is **not** in  $S$ and $c$ is a  nonzero constant, can $c\mathsf{v}$ be  in $S$?  


b. If $\mathsf{v}$ is **not** in $S$ and $\mathsf{w}$ is **not** in $S$, can  $\mathsf{v} + \mathsf{w}$ be in $S$? 


c. If $\mathsf{v}$ is in $S$ and $\mathsf{w}$ is **not** in $S$ can $\mathsf{v} + \mathsf{w}$ be in $S$?




## A Vector in Both Col(A) and Nul(A)

Give a $3 \times 3$ matrix $A$ for which the vector $\mathsf{v} = \begin{bmatrix}3 \\ -2 \\ 5   \end{bmatrix}$ is in both $\mathrm{Col}(A)$ and $\mathrm{Nul}(A)$. Be sure to demonstrate that $\mathsf{v} \in \mathrm{Col}(A)$ and $\mathsf{v} \in \mathrm{Nul}(A)$.



## Column and Null Space

Find a basis for the column space $Col(A)$ and the null space $Nul(A)$ of the matrix $A$ below
```{r,echo=TRUE}
(A = rbind(c(1, 2, 0, 2, 0, -1),c(1, 2, 1, 1, 0, -2),
           c(2, 4, -2, 6, 1, 2),c(1, 2,  0, 2, -1, -3 )))
```

## Shrink to a basis

Let $S$ be the subspace of $\mathbb{R}^5$ spanned by the vectors below. Find a basis for $S$ and find the dimension of $S$ (the number of vectors in a basis).

$$
\mathsf{v}_1 = \begin{bmatrix} 1\\ 1\\ 1\\ 1\\ 3 \end{bmatrix},
\mathsf{v}_2 = \begin{bmatrix} 1\\ -1\\ 5\\ 1\\ 1 \end{bmatrix},
\mathsf{v}_3 = \begin{bmatrix} 1\\ 1\\ 1\\ 0\\ 2 \end{bmatrix},
\mathsf{v}_4 = \begin{bmatrix} 2\\ -1\\ 8\\ 1\\ 2 \end{bmatrix},
\mathsf{v}_5 = \begin{bmatrix} 1\\ 1\\ 1\\ 1\\ 3 \end{bmatrix}.
$$
```{r,echo=TRUE}
(A = cbind(c(1,1,1,1,3),c(1,-1,5,1,1),c(1,1,1,0,2),c(2,-1,8,1,2),c(1,1,1,1,3)))
```

## Extend to a basis

I am  interested in the vectors below. I know that they do not span $\mathbb{R}^5$, because there are not enough of them, but I want to extend this set to a basis of $\mathbb{R}^5$ by adding some vectors to the set. 
$$
\begin{bmatrix} 5\\ 4\\ 3\\ 1\\ 2 \end{bmatrix},
\begin{bmatrix} 4\\ 4\\ 3\\ 1\\ 2 \end{bmatrix}, 
\begin{bmatrix} 1\\ 1\\ 1\\ 1\\ 1\end{bmatrix}. 
$$
I searched online for ideas and one suggested that I make the matrix below and row reduce it.
```{r,echo=TRUE}
(A = cbind(c(5,4,3,1,2),c(4,4,3,1,2),c(1,1,1,1,1),diag(5)))
```

a. Row reduce this matrix.

b. Use the result to come up with a basis for $\mathbb{R}^5$ that includes my original 3 vectors

c. Explain why this method works.


## 

Problem 8 is coming

