
# Problem Set 9

```{r}
require(pracma)
```


Download the [Rmd source file  for this problem set](https://raw.githubusercontent.com/Tom-Halverson/math236_s21/main/PS8.Rmd).


Upload a completed, knitted .html version of this file on Moodle. If you have collaborated with others on this assignment (encouraged), please include their names here (no penalty).

## Glucose-Insulin

The hormone insulin helps regulate glucose metabolism in your blood. The presene of insulin helps your body absorb excess glucose. Here $G_t$ (glucose) and $H_t$ (insulin) are measured as excess values (in mg per 100 ml of blood) above the steady state. 
$$
\begin{bmatrix} G_{t+1} \\ H_{t+1} \end{bmatrix}
=
\begin{bmatrix}0.9 & -0.4 \\0.1 & 0.9 \\\end{bmatrix}
\begin{bmatrix} G_t \\ H_t \end{bmatrix}= 
\begin{bmatrix} 0.9 G_t - 0.4 H_t \\ 0.1 G_t + 0.9 H_t \end{bmatrix}
$$

Here is what happens if we start at $(1,0)$ and iterate. That is we start with 1 unit excess glucose. Observe that the system spirals back to the steady state of $(0,0)$.
```{r,fig.width=6, fig.height=6}
A = cbind(c(.9,.1),c(-.4,.9))
start = c(1,0)
N = 50

X = matrix(0,nrow=2,ncol=N)
X[,1] = start

for (i in 2:N) {X[,i] = A %*% X[,i-1]}
plot(X[1,],X[2,],col='blue',xlim=c(-0.4,1.0),ylim=c(-0.4,0.4),xlab='glucose',ylab='insulin',pch=20,cex=1)
points(start[1],start[2],col="red")
grid(nx=10,ny=10)
```


We can plot the indivdual glucose and insulin coordinates over time. These are the x and y coordinates of the points in the above plot. You see the insulin responding to the excess glucose, and then the glucose being absorbed by the presence of insulin, and so on ...

```{r}
t = seq(1,N)
plot(t,X[1,],type='l',col='blue',ylim=c(-1,1),ylab="concentration (mg/ml)",xlab="time(hours)",main="Glucose-Insulin Over Time")
lines(t,X[2,],col='darkorange')
legend(10, 0.7, legend=c("Glucose", "Insulin"), col=c("blue", "darkorange"), lty=1)
```

The key point here is that the spiraling in the (x,y) plane or oscillating in the (x,t) plane corresponds to the presence of *complex eigenvalues*.  Your job is to perform an eigen-analysis of this problem:

a. Use R to find the eigenvalues and eigenvectors of $A$.

b. Write out the eigenvalues in the form $\lambda = a \pm b i$ and the eigenvectors in the form $\vec{v} = \vec{u} \pm \vec{w} i$ by filling in the values below:

$$
\begin{array}{lll}
\lambda_1 = a +  b i & \qquad & \lambda_2 = a -  b i \\ 
v_1 = \begin{bmatrix} xx \\ yy \end{bmatrix} + \begin{bmatrix} zz \\ ww \end{bmatrix}i &&
v_2 = \begin{bmatrix} xx \\ yy \end{bmatrix} - \begin{bmatrix} zz \\ ww \end{bmatrix}i
\end{array}
$$

c. Write the matrix $A$ in its rotation-dilationalized form by filling in the entries below with their appropriate values.

$$
\begin{bmatrix}0.9 & -0.4 \\0.1 & 0.9 \\\end{bmatrix}
= 
\begin{bmatrix} p_{11} & p_{12} \\ p_{21} & p_{22} \\\end{bmatrix}
\begin{bmatrix} a & b \\ -b & a \\\end{bmatrix}
\begin{bmatrix} p_{11}^{-1} & p_{12}^{-1} \\ p_{21}^{-1} & p_{22}^{-1} \\\end{bmatrix}
$$

d. Use R to multiply the matrices above and show that the really do combine to give $A$.

e. Find the scaling factor $|\lambda|$ for this matrix and the angle of rotation $\arctan(b/a)$. Give your answer in degrees. Compare your answers from part (c) to the plots above to confirm that the system is doing what the eigenvalues predict.


## The Rise of Moscow

Here is a video I made last year in Module 1 explaining the [Airline Network](https://macalester.voicethread.com/myvoice/thread/18172399) example.

If you collaborate on this one with classmates (and I hope you do), please include the names of the students that you worked with.

Read [Network Centralities] and analyze the network of trade routes in medieval Russia given to you there.

a. Create a vector containing the normalized Degree Centralities. See Section \@ref(degree-centrality) for help.

a. Create a vector containing the Gould Index values. See Section \@ref(gould-index-summary) for help.

a. Create a table that contains Gould's Index and Degree Centralities. The rows should be labeled with the city names and the columns should be named by the centrality measures. Sort according to Gould's Index.

a. Plot the network with the size of the vertices determined by Gould's Index and the size of the label is determined by degree centrality.

a. Use Gould's Index to decide whether Moscow's dominance was solely due to its geographic location.

a. Compare the Gould's Index and Degree Centrality rankings and note any interesting findings. See Section \@ref(step-5) for help.

## Orthogonal Complements

(Work on in class on Tuesday June 29)

Here are two subspaces of $\mathbb{R}^5$ that we have seen before. (See PS4.2 and PS5.2)
$$
\begin{align}
\mathsf{Z} & = \left\{ \begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \\ x_5 \end{bmatrix} \ \bigg\vert \ x_1 + x_2 + x_3 + x_4 + x_5 = 0,  x_4 = 2 x_2 \right\}.
\\
\mathsf{F} & = \left\{ \begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \\ x_5 \end{bmatrix} \ \bigg\vert \  x_3 =  x_1 + x_2, x_4 = x_2 + x_3, x_5 = x_3 + x_4 \right\}.
\end{align}
$$
Find the orthogonal complement of each subspace in $\mathbb{R}^5$. For each example, compute $\dim(W) + \dim(W^\perp)$.