# Problem Set 4



```{r}
require('pracma')
```












## Archaeological Seriation


There may be time to discuss this in class on Friday Sep 24.


The matrix $A$ below is used in  archaeological dating. Its rows correspond to four different grave sites $G_1, G_2, G_3, G_4$ and its columns correspond to five types of pottery$P_1, P_2, P_3, P_4, P_5$. There is a 1 in position $i$-$j$ if pottery type $P_j$ is found in grave $G_i$ (and a 0 otherwise).
$$
A=\begin{array}{c|ccccc}
& P_1 & P_2 & P_3 & P_4 & P_5 \\
\hline
 G_1 & 1 & 1 & 0 & 1 & 1 \\
G_2 &  0 & 0 & 1 & 0 & 1 \\
G_3 &  0 & 1 & 1 & 0 & 1 \\
G_4  & 1 & 1 & 1 & 1 & 1 \\
\end{array}
$$

a. Compute the matrix $\mathbf{G} = A A^T$, where $A^T$ is the *transpose* of $A$, meaning that the rows and columns have been interchanged.

b. Give the meaning of the $i$-$j$ entry of $\mathbf{G}$ (the entry in row $i$ and column $j$). State clearly the meaning of this entry using complete sentences (or sentence) and explain why it has this meaning.

## Rental Car

I hope to give you some time to work on this in class on Friday Sep 24.


Solve this problem using R and turn in a markdown file knitted to html. 

A group of Macalester alumni open a rental car company specializing in  renting electric cars. As a start, they have opened offices in St. Paul, Rochester, and Duluth. Through market research they find that of the cars rented in St. Paul, 85% will get returned in St. Paul, 9% will get returned in Rochester, and 6% will get returned in Duluth. Of the cars rented in Rochester, 30% will get returned in St. Paul, 60% will get returned in Rochester, and 10% in Duluth. Of the cars rented in Duluth, 35% will get returned in St. Paul, 5% in Rochester, and 60% in Duluth. This information is represented in the matrix below.
=======
The matrix $A$ below is used in  archaeological dating. Its rows correspond to four different grave sites $G_1, G_2, G_3, G_4$ and its columns correspond to five types of pottery$P_1, P_2, P_3, P_4, P_5$. There is a 1 in position $i$-$j$ if pottery type $P_j$ is found in grave $G_i$ (and a 0 otherwise).
$$
A=\begin{array}{c|ccccc}
& P_1 & P_2 & P_3 & P_4 & P_5 \\
\hline
 G_1 & 1 & 1 & 0 & 1 & 1 \\
G_2 &  0 & 0 & 1 & 0 & 1 \\
G_3 &  0 & 1 & 1 & 0 & 1 \\
G_4  & 1 & 1 & 1 & 1 & 1 \\
\end{array}
$$

1. Compute the matrix $\mathbf{G} = A A^T$, where $A^T$ is the *transpose* of $A$, meaning that the rows and columns have been interchanged.

2. Give the meaning of the $i$-$j$ entry of $\mathbf{G}$ (the entry in row $i$ and column $j$). State clearly the meaning of this entry using complete sentences (or sentence) and explain why it has this meaning.

## Rental Car

I hope to give you some time to work on this in class on Friday Sep 24.


```{r,echo=TRUE}
StP = c(.85,.09,.06)
Roch = c(.30,.60,.10)
Dul = c(.35,.05,.60) 
M = cbind(StP,Roch,Dul)
M
```
Such a matrix is called a *probability* matrix or a *stochastic* matrix because it  contains numbers between 0 and 1 and each of its columns sum to 1.

<<<<<<< HEAD
a. The owners are trying to use this data to estimate how much of their fleet will be at each location on average in the long run. Assume that initially they locate 20 cars in each city. This can be recorded by the vector `v0 = c(20,20,20)`. Apply, M to v0, call this vector v1, and explain, using how the matrix-vector product works, why v1 represents the number of cars at each location one day later (for simplicity, we assume that each rental is for 1 day).

b. Now apply M to v1 and call it v2. This should represent the number of cars at each location 2 days later. Also compute the square of the matrix M and call it M2. Confirm that M2 times v0 is the same as M times v1. 

c. Write a for loop that applies M over and over again to see what happens to the distribution of cars in the long-run (we will learn how to do this in class but you can also probably just google it). Does this sequence stabilize or does it keep changing after each application? If it does stabilize, how long does it take to stabilize (to within 0.1 cars at each location).

d. Does the starting distribution matter? Try 4 different starting distributions (with a total of 60 cars) and see what the final distribution looks like in each case. For one of your 4 starting distributions, try all 60 cars at one of the locations.
=======
Solve this problem using R and turn in a markdown file knitted to .html. 

A group of Macalester alumni open a rental car company specializing in  renting electric cars. As a start, they have opened offices in St. Paul, Rochester, and Duluth. Through market research they find that of the cars rented in St. Paul, 85% will get returned in St. Paul, 9% will get returned in Rochester, and 6% will get returned in Duluth. Of the cars rented in Rochester, 30% will get returned in St. Paul, 60% will get returned in Rochester, and 10% in Duluth. Of the cars rented in Duluth, 35% will get returned in St. Paul, 5% in Rochester, and 60% in Duluth. This information is represented in the matrix below.

```{r,echo=TRUE}
StP = c(.85,.09,.06)
Roch = c(.30,.60,.10)
Dul = c(.35,.05,.60) 
M = cbind(StP,Roch,Dul)
M
```
Such a matrix is called a *probability* matrix or a *stochastic* matrix because it  contains numbers between 0 and 1 and each of its columns sum to 1.

a. The owners are trying to use this data to estimate how much of their fleet will be at each location on average in the long run. Assume that initially they locate 20 cars in each city. This can be recorded by the vector `v0 = c(20,20,20)`. Apply, M to v0, call this vector v1, and explain, using how the matrix-vector product works, why v1 represents the number of cars at each location one day later (for simplicity, we assume that each rental is for 1 day).


b. Now apply M to v1 and call it v2. This should represent the number of cars at each location 2 days later. Also compute the square of the matrix M and call it M2. Confirm that M2 times v0 is the same as M times v1. 

<<<<<<< HEAD
## Rainy Day in LA

I hope to give you some time to discuss this in class on Monday Sep 24.


In Los Angeles if it rains today, there is a 50% chance it will rain tomorrow, but it if is sunny today, there is a 90% chance it will be sunny tomorrow. This is modeled in the rain-sunshine probability matrix P. 
$$
P = 
\begin{array}{c|cc|}
&\text{rain}&\text{sun}\\
\hline
\text{rain}&1/2&1/10\\
\text{sun}&1/2&9/10\\
\hline
\end{array}
$$
This matrix works as follows: if the rain-sunshine probability today is (40, 60) (that is, 40% chance rain and 60% chance sunshine), then the rain-sunshine probability tomorrow is (26, 74) as seen by the calculation below.
$$
\begin{bmatrix}
1/2 & 1/10 \\
1/2 & 9/10 \\
\end{bmatrix} \begin{bmatrix} 40 \\ 60 \end{bmatrix} = \begin{bmatrix} 26 \\ 74 \end{bmatrix}
$$

=======
c. Write a for loop that applies M over and over again to see what happens to the distribution of cars in the long-run (we will learn how to do this in class but you can also probably just google it). Does this sequence stabilize or does it keep changing after each application? If it does stabilize, how long does it take to stabilize (to within 0.1 cars at each location).

d. Does the starting distribution matter? Try 4 different starting distributions (with a total of 60 cars) and see what the final distribution looks like in each case. For one of your 4 starting distributions, try all 60 cars at one of the locations.


## Rainy Day in LA

I hope to give you some time to discuss this in class on Monday Sep 24.


In Los Angeles if it rains today, there is a 50% chance it will rain tomorrow, but it if is sunny today, there is a 90% chance it will be sunny tomorrow. This is modeled in the rain-sunshine probability matrix P. 
$$
P = 
\begin{array}{c|cc|}
&\text{rain}&\text{sun}\\
\hline
\text{rain}&1/2&1/10\\
\text{sun}&1/2&9/10\\
\hline
\end{array}
$$
This matrix works as follows: if the rain-sunshine probability today is (40, 60) (that is, 40% chance rain and 60% chance sunshine), then the rain-sunshine probability tomorrow is (26, 74) as seen by the calculation below.
$$
\begin{bmatrix}
1/2 & 1/10 \\
1/2 & 9/10 \\
\end{bmatrix} \begin{bmatrix} 40 \\ 60 \end{bmatrix} = \begin{bmatrix} 26 \\ 74 \end{bmatrix}
$$


a. Find the rain-sunshine probability the day after tomorrow.

b. Compute $P^2$ and explain the meaning of each of the four entries in the matrix.

c. Find $P^{-1}$ and   and use it find the rain-sunshine probability *yesterday* if the rain-sunshine probability today is (40, 60).



